---
title: "Untitled"
author: "chundra"
date: "5/16/2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(loo)
```

The models under comparison have the following order:

1. Predictors included, distributional
2. Predictors included, non-distributional
3. Predictors not included, distributional
4. Predictors not included, non-distributional

```{r}
loo.objects <- readRDS('model.comparisons.RDS')

dominant.compare <- loo.objects[[1]]
dominant.stack <- loo.objects[[2]]
rhythm.compare <- loo.objects[[3]]
rhythm.stack <- loo.objects[[4]]
```

# Dominant frequency

In terms of stacking weight, we see that the two distributional models perform roughly equally. 

```{r}
dominant.stack
```

This means that including predictors such as weight of dominant frequency does not greeatly improve model predictions. This is a bit odd. Perhaps the random effects structure is too conservative?

# Rhythm

In terms of stacking weight, we see that the null distributional models greatly performs the distributional model with predictors. 

```{r}
rhythm.stack
```
