---
title: "Untitled"
author: "chundra"
date: "5/16/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(loo)
require(knitr)
```

The models under comparison have the following order:

1. Predictors included, distributional
2. Predictors included, non-distributional
3. Predictors not included, distributional
4. Predictors not included, non-distributional

```{r}
loo.objects <- readRDS('model.comparisons.RDS')

dominant.compare <- loo.objects[[1]]
dominant.stack <- loo.objects[[2]]
rhythm.compare <- loo.objects[[3]]
rhythm.stack <- loo.objects[[4]]
```

# Raw ELPD values

## Dominant frequency

```{r}
kable(dominant.compare)
```

The distributional model with predictors included has the highest ELPD value, but it is not substantially higher than the null distributional model (e.g., more than 2 std. errors higher).

## Rhythm

```{r}
kable(rhythm.compare)
```

The null distributional model has the highest ELPD value, but it is not substantially higher than the distributional model with predictors (e.g., more than 2 std. errors higher).

# Stack

## Dominant frequency

In terms of stacking weight, we see that the distributional models with predictors has the majority of the stacking weight, but the null distributional model still has a substantial amount of stacking weight. 

```{r}
dominant.stack
```

## Rhythm

In terms of stacking weight, we see that the null distributional model outperforms the distributional model with predictors but the full distributional model has a not insignificant portion of the stacking weight. 

```{r}
rhythm.stack
```